Технология использования открытых данных Федеральной налоговой службы России для экономико-географического анализа

### 2024-07-06 law-firms-geography/Paper.Rmd
Rename folder, #1
Edit according to reviewer's notes: add custom definition of law business in methods, use per capita counts in Fig. 1 instead of absolute ones, #1

The research on the geography of legal services in Russia is limited. To date, only an analytical review describing the legal market in Russia has been published, which contains a chapter on its spatial aspects [@Moiseeva2016]. Another paper briefly addresses the role of Russia in the global legal services market without delving into the internal spatial properties of the Russian legal business [@Yastrebov2021]. In addition to scientific publications, there are business products about the legal services market in Russia that include its regional analysis[^2]. Despite a low observed interest in this issue from other researchers, it can be argued that geographical studies of legal services are essential. First, legal services, as a subset of other business or producer services, play a substantial role in the development of regions and territories and economic growth [@Hansen1993]. In some cases, the lack of professional legal advice may even act as a limiting factor, causing problems for entrepreneurs, especially small ones, and decreasing their performance [@Hitchens1997]. Next, a comprehensive examination of law firms may yield valuable insights into the business sectors they serve [@Faulconbridge2019] or enhance the theoretical understanding of markets and industries in general [@Henderson2016]. Finally, legal services should be considered not only in terms of their narrow economic function but also as an integral element of the broader legal system. This system is responsible for protecting human rights and freedoms, ensuring the rule of law, and acting as an intermediary between people, businesses, and government. It helps to resolve conflicts, increase certainty and orderliness in social relations and thus contributes to the stability and development of society. This aspect of legal services is closely related to the social dimension of the geography of services. It is therefore unsurprising that research is being conducted on the problem of legal services availability [@Baxter2015].

[^2]: Analysis of legal services market in Russia in 2019–2023 with forecast for 2024–2028 (in Russian). Web resource: https://marketing.rbc.ru/research/41683/ (accessed 14.04.2024)

@@

The SMB registry dataset contains full dumps of the respective registry in zipped XML files. The initial dataset was made available in August 2016, and updates have been released every month. Each dump contains the entire registry and includes, in particular, the company or individual names, taxpayer IDs, and activity codes according to the All-Russia Classifier of Kinds of Economic Activity, as well as addresses of incorporation (up to the settlement name). According to the Russian laws, the registry includes information about all commercial organizations and individual entrepreneurs that meet several criteria (SMB subjects). The most important criteria are limits on the revenue (no more than 2B Russian rubles) and on the average number of employees (no more than 250). Consequently, the open dumps of this registry represent the most comprehensive publicly accessible data source on the Russian business entities. Moreover, the archive of these dumps constitutes a time series spanning from August 2016 to the present, thereby providing a history of small and medium-sized business changes.

The Information about Employees dataset contains company taxpayer IDs and average yearly numbers of employees for the previous year. The data was initially published in 2019 and has been updated with varying frequencies, though at least annually. Consequently, the data archive encompasses the period from 2018 to 2023. For reasons that are not specified, the older versions of the dataset which included information for the years from 2018 to 2022 are not currently available on the FTS website. However, the data had already been downloaded by the author before its removal, and thus it is used in this research.

@@

Apart from these datasets, a small manually collected table with the locations of Russian district-level courts is utilized. District-level courts span their activity over two or more regions of Russia that constitute a judicial district, checking and reviewing decisions of lower-level courts. These category include appeal and cassation ordinary and commercial courts, with a total of 47 such courts in Russia. The location data was obtained from the Supreme Court of Russia website[^7] and the legal information system called “Consultant Plus”[^8]. While not the primary focus of this paper, the court location data was utilized to supplement the regional analysis by examining the relationship between the legal business geography and the placement of district-level courts.

@@

Finally, an aggregated official statistic on the annual total number of employees in the regions of Russia[^9] was utilized. This data is provided by the Federal Statistics Service (Rosstat) through the publicly available Unified Interdepartmental Information and Statistical System. It contains region names and corresponding numbers of employees in all sectors of the economy for each selected year.

[^9]: Number of people employed (in Russian). Web resource: https://fedstat.ru/indicator/34051 (accessed 24.04.2024)

**Methods.** A Python command line tool, ru-smb-companies, was developed by the author of this paper to initially process large and complex raw XML datasets from the FTS servers and transform them into a single CSV table. The source code of this tool, along with a brief documentation, is published in the author's personal GitHub account[^10]. Here, a concise summary of the data processing workflow is provided. The data processing takes several steps.

[^10]: Russian small and medium business dataset generator. Web resource: https://github.com/PavelSyomin/russian-smb-companies (accessed 15.04.2024)

@@

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “law firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “lawyers” in this paper refers to the employees of these firms.

Problem 1 is the distribution of law firms across Russia's regions. To discover this distribution, the median number of law firms in each region is calculated for the period from 2016 to 2021. This result is then visualized on a map.

@@

Technically, the analysis is carried out using the R programming language (including additional packages). A tile grid map of Russia by I. Dianov[^11] is used with a modified implementation by Yu. Tukachev[^12]. The source code of this paper is published in the author's repository on GitHub[^13].

[^11]: Tile grid map of Russia (in Russian). Web resource: https://dianov.org/all/plitochnaya-karta-rossii/ (accessed 23.04.2024)
[^12]: Tile grid map of Russia. Web resource: https://gist.github.com/tukachev/c3262242a20b3050e0459b4e3afc3a51 (accessed 23.04.2024)
[^13]: PavelSyomin / ru-smb-companies-papers. Web resource: https://github.com/PavelSyomin/ru-smb-companies-papers/tree/main/legal-companies-geography (accessed 23.04.2024)

@@

Fig. 1 depicts the average annual number of law firms by region. Several conclusions can be drawn from this visualization. First, Moscow is the only city where the number of law firms exceeds 10 thousand, and this is the highest value in Russia. Second, some regions also have a relatively high number of law firms. Geographically, these regions form several groups. One of these groups is located in the south of the European part of Russia and consists of Krasnodar and Stavropol territories with Rostov, Volgograd, and Voronezh regions. The Republic of Bashkortostan and the Republic of Tatarstan with Samara, Sverdlovsk, and Chelyabinsk regions, as well as Perm Territory, form the Ural-Volga group. Tyumen, Omsk, Novosibirsk, Kemerovo regions with Krasnoyarsk Territory belong to the third group which can be designated as Siberian due to its location. The city of St. Petersburg, Nizhniy Novgorod Region, and Primorye Territory are separate regions with a relatively high number of law firms. Third, the majority of other regions have a relatively small number of law firms, with several republics of the North Caucasus and Altai and a portion of regions of the Russian North incorporating less than a hundred of such companies. It can be hypothesized that the number of law firms is positively associated with the economic development of a region. Furthermore, the population of a region may also be a factor to consider.

The relative location of regions and groups with different numbers of law firms suggests the possibility of central-peripheries relations between them. For instance, Moscow City and Region, St. Petersburg City, and Nizhny Novgorod Region may serve as hubs for their surrounding regions of Central and North-Western Russia. A similar phenomenon may be observed in other parts of the country: the Siberian group provides legal services for the Far East, while the regions of the European South serve the republics of the North Caucasus.

@@

```{r fig-2-shares-map, fig.cap="Fig. 2. The median annual share of companies among legal firms by regions of Russia over the period from 2016 to 2021"}
shares_map
```

@@

The changes in the number of law firms by region are depicted in Fig. 4. The number of such firms increased in the majority of regions, indicating the development of the legal business sector. In contrast, several regions, all of which are located in the European part of Russia, have experienced a net decrease in the number of law firms, In comparison to the increase rates observed in other regions, the net change in the number of law firms from 2016 to 2021 in these regions is relatively low (usually less than 10%). In some cases, neither decline nor growth was clearly expressed (for example, the Republic of Altai, Lipetsk, and Tambov regions). The regions with the most significant growth are some republics in the North Caucasus, as well as the Republic of Kalmykia and Nenets Autonomous Area. Their results are likely attributable to the low base effect, given that the absolute counts of law firms in all these regions are below 100 (see Fig. 1). It is noteworthy that Leningrad and Moscow regions also demonstrated one of the highest growth rates. It is likely that the relocation of law firms from the cities of Moscow and St. Petersburg to the surrounding regions has contributed to this outcome. Nevertheless, both Moscow and St. Petersburg cities also exhibited high net gain of law firms.

@@

The regional analysis revealed several spatial features of legal business in Russia. First, the regions exhibited considerable variation in the number of law firms and lawyers. Several regional clusters with relatively high development of legal business were identified, including the European South, Volga-Ural, and Siberian. As might be expected, the cities of Moscow and St. Petersburg are home to the largest number of law firms and lawyers. Furthermore, it is again Moscow and St. Petersburg cities that have the highest share of lawyers in their total labor force. Nevertheless, there are other regions with a relatively high role of legal businesses in their regional economies, including Sverdlovsk, Tyumen, and Novosibirsk.

>>>>

The research on the geography of legal services in Russia is limited. To date, only an analytical review describing the legal market in Russia has been published, which contains a chapter on its spatial aspects [@Moiseeva2016]. Another paper briefly addresses the role of Russia in the global legal services market without delving into the internal spatial properties of the Russian legal business [@Yastrebov2021]. In addition to scientific publications, there are business products about the legal services market in Russia that include its regional analysis[^2]. Despite a low observed interest in this issue from other researchers, it can be argued that geographical studies of legal services are essential. First, legal services, as a subset of other business or producer services, play a substantial role in the development of regions and territories and economic growth [@Hansen1993]. In some cases, the lack of professional legal advice may even act as a limiting factor, causing problems for entrepreneurs, especially small ones, and decreasing their performance [@Hitchens1997]. Next, a comprehensive examination of law firms may yield valuable insights into the business sectors they serve [@Faulconbridge2019] or enhance the theoretical understanding of markets and industries in general [@Henderson2016]. Finally, legal services should be considered not only in terms of their narrow economic function but also as an integral element of the broader legal system. This system is responsible for protecting human rights and freedoms, ensuring the rule of law, and acting as an intermediary between people, businesses, and government. It helps to resolve conflicts, increase certainty and orderliness in social relations, and thus contributes to the stability and development of society. This aspect of legal services is closely related to the social dimension of the geography of services. It is therefore unsurprising that research is being conducted on the problem of legal services availability [@Baxter2015].

[^2]: Analysis of the legal services market in Russia in 2019–2023 with forecast for 2024–2028 (in Russian). Web resource: https://marketing.rbc.ru/research/41683/ (accessed 14.04.2024)

@@

The SMB registry dataset contains full dumps of the respective registry in zipped XML files. The initial dataset was made available in August 2016, and updates have been released every month. Each dump contains the entire registry and includes, in particular, the company or individual names, taxpayer IDs, and activity codes according to the All-Russia Classifier of Kinds of Economic Activity, as well as addresses of incorporation (up to the settlement name). According to Russian laws, the registry includes information about all commercial organizations and individual entrepreneurs that meet several criteria (SMB subjects). The most important criteria are limits on the revenue (no more than 2B Russian rubles) and on the average number of employees (no more than 250). Consequently, the open dumps of this registry represent the most comprehensive publicly accessible data source on Russian business entities. Moreover, the archive of these dumps constitutes a time series spanning from August 2016 to the present, thereby providing a history of small and medium-sized business changes.

The Information about employees dataset contains company taxpayer IDs and average yearly numbers of employees for the previous year. The data was initially published in 2019 and has been updated with varying frequencies, though at least annually. Consequently, the data archive encompasses the period from 2018 to 2023. For reasons that are not specified, the older versions of the dataset which included information for the years from 2018 to 2022 are not currently available on the FTS website. However, the data had already been downloaded by the author before its removal, and thus it is used in this research.

@@

Apart from these datasets, a small manually collected table with the locations of Russian district-level courts is utilized. District-level courts span their activity over two or more regions of Russia that constitute a judicial district, checking and reviewing decisions of lower-level courts. This category include appeal and cassation ordinary and commercial courts, with a total of 47 such courts in Russia. The location data was obtained from the Supreme Court of Russia website[^7] and the legal information system called “Consultant Plus”[^8]. While not the primary focus of this paper, the court location data was utilized to supplement the regional analysis by examining the relationship between the legal business geography and the placement of district-level courts.

@@

Finally, an aggregated official statistic on the annual total number of employees in the regions of Russia[^9] and on the average annual resident population[^10] was utilized. This data is provided by the Federal Statistics Service (Rosstat) through the publicly available Unified Interdepartmental Information and Statistical System. It contains region names and corresponding numbers of employees in all sectors of the economy for each selected year.

[^9]: Number of people employed (in Russian). Web resource: https://fedstat.ru/indicator/34051 (accessed 24.04.2024)
[^10]: Average annual resident population (in Russian). Web resource: https://fedstat.ru/indicator/31556 (accessed 24.04.2024)

**Methods.** A Python command line tool, ru-smb-companies, was developed by the author of this paper to initially process large and complex raw XML datasets from the FTS servers and transform them into a single CSV table. The source code of this tool, along with a brief documentation, is published in the author's personal GitHub account[^11]. Here, a concise summary of the data processing workflow is provided. The data processing takes several steps.

[^11]: Russian small and medium business dataset generator. Web resource: https://github.com/PavelSyomin/russian-smb-companies (accessed 15.04.2024)

@@

The research adheres to the general framework of regional analysis and employs the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. In this study, the concept of legal business is approached in a relatively narrow sense and with a strong emphasis on formal considerations. The term encompasses (1) commercial legal entities or individual entrepreneurs (2) registered in Russia according to the Russian civil law (3) whose main activity is classified under code 69.10 in the All-Russia Classifier of Kinds of Economic Activity. This technique is consistent with the selection criterion for law firms used in [@Moiseeva2016]. The analysis does not include non-profit legal services providers or specialized professionals such as notaries and advocates. Due to data availability constraints, it also omits large companies that are not present in the small and medium-sized businesses registry. In essence, the legal business, as it is defined in this context, encompasses legal counseling or similar activities. The analysis is comprised of four research problems, or questions, which are enumerated below with the description of the relevant techniques used to find the answers. The term “law firm” denotes both companies (organizations) and individuals (individual entrepreneurs). The term “lawyers” in this paper refers to the employees of these firms.

Problem 1 is the distribution of law firms across Russia's regions. To discover this distribution, the median number of law firms in each region per 100,000 population is calculated for the period from 2016 to 2021. This result is then visualized on a map.

@@

Technically, the analysis is carried out using the R programming language (including additional packages). A tile grid map of Russia by I. Dianov[^12] is used with a modified implementation by Yu. Tukachev[^13]. The source code of this paper is published in the author's repository on GitHub[^14].

[^12]: Tile grid map of Russia (in Russian). Web resource: https://dianov.org/all/plitochnaya-karta-rossii/ (accessed 23.04.2024)
[^13]: Tile grid map of Russia. Web resource: https://gist.github.com/tukachev/c3262242a20b3050e0459b4e3afc3a51 (accessed 23.04.2024)
[^14]: PavelSyomin / ru-smb-companies-papers. Web resource: https://github.com/PavelSyomin/ru-smb-companies-papers/tree/main/law-firms-geography (accessed 23.04.2024)

@@

Fig. 1 depicts the average relative annual number of law firms by region. Several conclusions can be drawn from this visualization. First, Moscow and Saint Petersburg stand out as the only cities in Russia where the number of law firms exceeds 100 per 100,000 inhabitants. Secondly, certain regions also have a relatively high number of law firms per capita. Geographically, these regions are situated in different parts of the country, and in some cases, they form clusters. The most prominent of these groups is located in the Middle and Southern Urals with Western Siberia and consists of Sverdlovsk, Chelyabinsk, Tyumen, Omsk, and Novosibirsk regions. Another group is located in the Far East and includes Primorye territory, Sakhalin region, and, somewhat surprisingly, Kamchatka territory. Krasnodar territory, the Republic of Tatarstan, the Republic of Karelia, and Arkhangelsk region are distinct centers with a high number of law firms per capita surrounded by the regions with a more moderate number of such firms. Moscow region and its neighboring Ivanovo region are also characterized by a considerable number of law firms per capita, but it seems that their metrics may be the result of a “spillover effect” originating from Moscow city. Third, the majority of other regions have a relatively small number of law firms, with several republics of the North Caucasus and Altai and a portion of regions of the Russian North, sometimes incorporating less than ten such companies. It can be hypothesized that the number of law firms per capita is contingent upon the economic development of a region.

The relative location of regions and groups with different numbers of law firms suggests the possibility of central-periphery relations between them. For instance, Moscow City and Region, St. Petersburg City, and the Republic of Tatarstan may serve as hubs for their surrounding areas of Central, North-Western Russia, and Volga region (Povolzhye). A similar phenomenon may be observed in other parts of the country: the Ural-Siberian cluster provides legal services for Eastern Siberia and the Far East, while the Krasnodar Territory serves the Rostov and Stavropol.

@@

```{r fig-2-shares-map, fig.cap="Fig. 2. The median annual share of companies among law firms by regions of Russia over the period from 2016 to 2021"}
shares_map
```

@@

The changes in the number of law firms by region are depicted in Fig. 4. The number of such firms increased in the majority of regions, indicating the development of the legal business sector. In contrast, several regions, all of which are located in the European part of Russia, have experienced a net decrease in the number of law firms, In comparison to the increase rates observed in other regions, the net change in the number of law firms from 2016 to 2021 in these regions is relatively low (usually less than 10%). In some cases, neither decline nor growth was clearly expressed (for example, in the Republic of Altai, Lipetsk, and Tambov regions). The regions with the most significant growth are some republics in the North Caucasus, as well as the Republic of Kalmykia and Nenets Autonomous Area. Their results are likely attributable to the low base effect, given that the absolute counts of law firms in all these regions are below 100 (see Fig. 1). It is noteworthy that Leningrad and Moscow regions also demonstrated one of the highest growth rates. It is likely that the relocation of law firms from the cities of Moscow and St. Petersburg to the surrounding regions has contributed to this outcome. Nevertheless, both Moscow and St. Petersburg cities also exhibited high net gain of law firms.

@@

The regional analysis revealed several spatial features of legal business in Russia. First, the regions exhibited considerable variation in the number of law firms and lawyers. Several areas with a relatively high number of lawyers per capita were identified, including the Urals and Western Siberia cluster, a few regions of the Far East, Krasnodar territory, the Republic of Tatarstan, and a number of other regions. As might be expected, the cities of Moscow and St. Petersburg are home to the largest number of law firms and lawyers. Furthermore, it is again Moscow and St. Petersburg cities that have the highest share of lawyers in their total labor force. Nevertheless, there are other regions with a relatively high role of legal businesses in their regional economies, including Sverdlovsk, Tyumen, and Novosibirsk.
